Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-07-12 MCQMC2022ChallengesQMCSoftwareArticle_revision_1.tex
Updates from Overleaf

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{UMBridge23}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$, and their derivatives if applicable. UM-Bridge allows coupling of UQ codes to models regardless of the programming languages.
{eq:cubSummary}
We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam of length~$L$, with a random Young's modulus $m(x)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and
the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. }
%approximates $u(x)$ and $m(x)$ at the nodes, respectively.
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$.
The beam model was written in C++ using the MUQ's model framework~\cite{MUQ2}. It is one of the UM-Bridge's benchmark problem library, available as a container.

@@

In spite of the desire for efficiency, great QMC software will be written in a variety of computer languages since different communities tend to favor different languages. Trade-offs will be made between efficiency and convenience.
For example, as noted by a referee, QMCPy is not particularly fast at generating tens of thousands of replicates of randomized QMC sample means in comparison to SSJ, which is based on Java. Such cases may point to the need for a Python layer over a faster implementation.

@@

Python is known for its versatility, but it can be slower than highly optimized implementations in languages like Java or C++.  For QMCPy developers, it would be beneficial to investigate such underlying interfaces or consider other code acceleration techniques by leveraging high-performance libraries such as Numba, JAX, Joblib, or Ray in the future.

>>>>

Complex integrands might be best evaluated by libraries that have no direct connection to QMC software libraries.  In the uncertainty quantification (UQ) space, the recently developed UQ and Model Bridge (UM-Bridge) \cite{UMBridge23}, connects libraries that generate sequences of model parameters, $\bsx_1, \bsx_2, \ldots$ with libraries that evaluate models, $f(\bsx_1), f(\bsx_2), \ldots$, and their derivatives if applicable. UM-Bridge allows coupling of UQ codes to models regardless of the programming languages.
{eq:cubSummary}
We illustrate in Figure~\ref{CDHJSfigbeam} how UM-Bridge connects QMCPy with a differential equation boundary value problem solver for a cantilevered beam of length~$L$, with a random Young's modulus $m(x)$, assumed to be piecewise constant over three regions, each of which with a uniform distribution of uncertainty~\cite{ParSee22a}, i.e., $m(x) = \left[\begin{matrix} m_1(x), m_2(x), m_3(x) \end{matrix} \right]$ with $m_i(x) \sim \calU[1,1.2]^3$.  The beam position $x$ is discretized at $N=31$ equidistant points, and
the vertical deflection $u(x)$ when the beam is subject to a given external load  $g(x)$ satisfies the Euler-Bernoulli partial differential equation (PDE):
$$ \frac{\partial^2}{\partial x^2}\left[ \exp(m(x)) \frac{\partial^2  u(x)}{\partial x^2}\right] = g(x),$$
%where $m(x) = \log E(x)$ is the log of an effective stiffness $E(x)$.  The
with the following boundary conditions:
$$ \left.u(x)\right|_{x=0} = 0, \qquad
\left.\frac{\partial u(x)}{\partial x}\right|_{x=0} = 0, \qquad
\left.\frac{\partial^2 u(x)}{\partial x^2}\right|_{x=L} = 0, \qquad  \left.\frac{\partial^3 u(x)}{\partial x^3}\right|_{x=L} = 0. $$
This PDE can be discretized with finite differences yielding a linear system $K(m) \hat{u} = \hat{g}$, where $\hat{u}$ and $\hat{g} \in \mathbb{R}^N$. 
%approximates $u(x)$ and $m(x)$ at the nodes, respectively.
%The forward model $F:\mathbb{R}^3 \rightarrow \mathbb{R}^N $ maps $m = \left[\begin{matrix}m_1, m_2, m_3 \end{matrix}\right]^T$ to $\hat{u}$, while $Q : \mathbb{R}^N \rightarrow \mathbb{R}^2 $ selects the estimated solution at the 10th and 25th nodes, i.e., $ Q(\hat{u}) := \left[\begin{matrix} \hat{u}_{10}, \hat{u}_{25} \end{matrix} \right]^T$. The objective is to determine the distributions of $Q(F(m))$.
The beam model was written in C++ using the MUQ's model framework~\cite{MUQ2}. It is one of the UM-Bridge's benchmark problem library, available as a container.

@@

In spite of the desire for efficiency, great QMC software will be written in a variety of computer languages since different communities tend to favor different languages. Trade-offs will be made between efficiency and convenience.
As noted by a referee, QMCPy is not particularly fast at generating tens of thousands of replicates of randomized QMC sample means in comparison to SSJ, which is based on Java. Such cases may point to the need for a Python layer over a faster implementation.

@@

Although Python is popular and versatile, but it can be slower than highly optimized implementations in languages like Java or C++.  Developers of QMC software in Python  developers would do well to investigate interfaces to optimized code in other languages or consider other code acceleration techniques provided by libraries such as Numba, JAX, Joblib, or Ray.  With time, languages rise and fall in popularity, and those who wish to provide great QMC software may need to pivot to new software environments.
